### 2018-08-02 VV570/PROJ2/tex/report.tex
add task3 report

\section{Task 3}
	\begin{enumerate}[(a)]
		\item

@@

@@

>>>>

\section{Task 3}
	\begin{enumerate}[(a)]
		\item 
			Denote $y_1(t) \triangleq n(t)$, $y_2(t) \triangleq x(t)$ and
			\begin{equation}
				\Phi \triangleq
				\begin{bmatrix}
					\dot{y_1}\\
					 \dot{y_2}	
				\end{bmatrix}
				=
				\begin{bmatrix}
					\gamma - \alpha y_2^2 y_1\\
					\alpha y_2^2 y_1 - \frac{y_1}{1+y_1}
				\end{bmatrix}
			\end{equation}
			Then, the synthetic data for $n$ and $x$ can be calculated iteratively. To make it more accurate, the Runge-Kutta 4-stage method was adopted. The matlab built-in function 'ode45' was used in my implementation. \\
			Plot of $n(t)$ and $x(t)$ were generated as follows
			\begin{figure}[htbp]\centering
				\begin{minipage}{7.5cm}
					\centering
					\includegraphics[height=7cm,width=7cm]{../pic/n_and_ns.eps}
				\end{minipage}%
				\begin{minipage}{7.5cm}
					\centering
					\includegraphics[height=7cm,width=7cm]{../pic/x_and_xs.eps}
				\end{minipage}
			\end{figure}

			Plot of $n(t)$ and $x(t)$ with noise were also generated here to make a clear comparsion.
		\item 
			At each point (except the two boundary points), the 1st-order derivative of $n(t)$ and $x(t)$ can be approximated using central difference as
			\begin{equation}
				\dot{n_k} = \frac{n_{k+1}-n_{k-1}}{2 \Delta t} \label{si_eq1}
			\end{equation}
			\begin{equation}
				\dot{x_k} = \frac{x_{k+1}-x_{k-1}}{2 \Delta t} \label{si_eq2}
			\end{equation}
			Thus a pair of $\alpha$ and $\gamma$ can be solved at each point as
			\begin{equation}
				\alpha_k = \frac{\dot{x_k} + \frac{x_k}{1+x_k}}{n_k^2x_k}
			\end{equation}
			\begin{equation}
				\gamma_k = \dot{x_k} + \dot{n_k} + \frac{x_k}{1+x_k}
			\end{equation}
			for $k=2,3,...,N-1$.\\
			Finally, the estimated $\alpha$ and $\gamma$ can be calculated as
			\begin{equation}
				\hat{\alpha} = \frac{\sum_{k=2}^{N-1}\alpha_k}{N-2}
			\end{equation}
			\begin{equation}
				\hat{\gamma} = \frac{\sum_{k=2}^{N-1}\gamma_k}{N-2}
			\end{equation}
			Numerical experiment shows that $\hat{\alpha} = 0.099824$ and $\hat{\gamma} = 0.199699$, which implies that $\hat{\alpha}$ and $\hat{\gamma}$ agree well with $\alpha$ and $\gamma$.
		\item 
			The noise is generated with the built-in function 'normrnd' of matlab in my code, plots are given in part(a).

@@

			Considering the mathmatical expect of $Eq(\ref{si_eq1})$ and $Eq(\ref{si_eq2})$ as
			\begin{equation}
				\begin{aligned}
					E(\dot{n}) & = \gamma - \alpha E(n^2 x)\\
					E(\dot{x}) & = \alpha E(n^2 x) - E(\frac{x}{1+x})
				\end{aligned}
			\end{equation}
			Then $\alpha$ and $\gamma$ can be resolved as
			\begin{equation}
				\begin{aligned}
					\hat{\alpha} & = \frac{E(\dot{x}) + E(\frac{x}{1+x})}{E(n^2 x)}\\
					\hat{\gamma} & = E(\dot{x}) + E(\dot{n}) + E(\frac{x}{1+x})
				\end{aligned}
			\end{equation}
			But the signal we have is not clean, what we can obtain from the data are quantities with noise like $E(\dot{n^*})$, $E(\dot{x^*})$, $E({n^*}^2 x^*)$, and $E(\frac{x^*}{1+x^*})$. These clean quantities like $E(\dot{n})$, $E(\dot{x})$, $E(n^2 x)$, and $E(\frac{x}{1+x})$ should be approximated from these polluted quantities so that $\hat{\alpha}$ and $\hat{\gamma}$ can be resolved.\\
			As $\epsilon \sim N(0, 0.01)$ and $\nu \sim N(0, 0.001)$, then $E(\epsilon) = 0$, $E(\nu) = 0$ and $E(\epsilon^2) = 0.01$, $E(\nu^2)=0.001$.
			Since $n^* = n + \epsilon$ and $x^* = x + \nu$, then
			\begin{equation}
				\begin{aligned}
					E(n^*) & = E(n) + E(\epsilon) = E(n)\\
					E(x^*) & = E(x) + E(\nu) = E(x)
				\end{aligned}
			\end{equation}
			As $\epsilon$ and $\nu$ are independent, thus
			\begin{equation}
				\begin{aligned}
					E({n^*}^2 x^*) & = E(n^2 x + 2nx\epsilon + x\epsilon^2 + n^2\nu + 2n\epsilon\nu + \epsilon^2 \nu)\\
					               & = E(n^2 x) + E(x)E(\epsilon^2)
				\end{aligned}
			\end{equation}
			and
			\begin{equation}
				E(\frac{x^*}{1+x^*}) = E(\frac{x}{1+x^*}) \approxeq E(\frac{x}{1+x})
			\end{equation}
			For derivatives, similiar relation can be concluded as
			\begin{equation}
				\begin{aligned}
					E(\dot{n_k^*}) & = \frac{E(n_{k+1}^*) - E(n_{k-1}^*)}{2 \Delta t} = \frac{E(n_{k+1}) - E(n_{k-1})}{2 \Delta t} = E(\dot{n_k})\\
					E(\dot{x_k^*}) & = \frac{E(x_{k+1}^*) - E(x_{k-1}^*)}{2 \Delta t} = \frac{E(x_{k+1}) - E(x_{k-1})}{2 \Delta t} = E(\dot{x_k})
				\end{aligned}
			\end{equation}
			Hence these clean quantities can be approximated as
			\begin{equation}
				\begin{aligned}
					E(n) & = E(n^*)\\
					E(x) & = E(x^*)\\
					E(\dot{n}) & = E(\dot{n^*}) = \frac{\sum_{k=2}^{N-1}\dot{n_k^*}}{N-2}\\
					E(\dot{x}) & = E(\dot{x^*}) = \frac{\sum_{k=2}^{N-1}\dot{x_k^*}}{N-2}\\
					E(n^2 x) & = E({n^*}^2 x^*) - E(x)E(\epsilon^2)\\
					E(\frac{x}{1+x}) & = E(\frac{x^*}{1+x^*})
				\end{aligned}
			\end{equation}
			where the properties of the noise is utilized.\\
			In this way, $\hat{\alpha}$ and $\hat{\gamma}$ were resolved in numerical experiment as
			\begin{equation}
				\begin{aligned}
					\hat{\alpha} & = 0.102744\\
					\hat{\gamma} & = 0.201082
				\end{aligned}
			\end{equation}
			The result has been validated for about 20 times, although the noise is different at each time, as well as $\hat{\alpha}$ and $\hat{\gamma}$, but each time the result was quite stable.

@@

			If the properties of noise are not available, some approximations or assumptions can not be made finely.\\
			At first, a distinct attempt is to treat $n^*$ and $x^*$ like what we have done in part(b).
			\begin{equation}
				\hat{\alpha} = \frac{\sum_{k=2}^{N-1}\alpha_k^*}{N-2}
			\end{equation}
			\begin{equation}
				\hat{\gamma} = \frac{\sum_{k=2}^{N-1}\gamma_k^*}{N-2}
			\end{equation}
			In this way, the only difference is the input data sequence, as the input in part(b) is clean but here is polluted. \\
			Numerical experiments were done for 200 times to illustrate the initial guess(Fig\ref{coarse_ag}) and the variance(Fig\ref{err_ag}) of $\hat{\alpha}$ and $\hat{\gamma}$.
			\begin{figure}[!htbp]\centering
				\includegraphics[height=10cm]{../pic/coarse_estimation.eps}
				\caption{Coarse estimation} 
				\label{coarse_ag}
			\end{figure}
			\begin{figure}[!htbp]\centering
				\includegraphics[height=10cm]{../pic/coarse_estimation_err.eps}
				\caption{Error of estimation}
				\label{err_ag}
			\end{figure}

			It can be seen from Fig\ref{err_ag} that the error of $\hat{\gamma}$ is about 3\% relative to $\bar{\gamma}$, which is quite nice. But the error of $\hat{\alpha}$ is about 200\% relative to $\bar{\alpha}$, which means that the estimation of $\alpha$ fails in this manner.
			Hence, more sophisticated estimation need to be adopted, although this coarse approach gives relative precise estimation of $\gamma$.\\
			Considering the approach in part(e), where $E(n^2 x)$ is estimated utilizing the info of the noise, similiar approximations can be made as
			\begin{equation}
				E(n^2 x) = E({n^*}^2 x^*) = \frac{\sum_{k=1}^{N}{n_k^*}^2 x_k^*}{N}
			\end{equation}
			and
			\begin{equation}
				E(\dot{n}) = E(\dot{n^*}) = \frac{\sum_{k=2}^{N-1}\dot{n_k^*}}{N-2}
			\end{equation}
			Here, no assumption of the distribution of the noise was made. With $\hat{\gamma}$ given by the coarse estimation, a relative finer estimation of $\hat{\alpha}$ can be calculated as
			\begin{equation}
				\hat{\alpha} = \frac{\hat{\gamma} - E(\dot{n^*})}{E({n^*}^2 x^*)}
			\end{equation} 
			Numerical experiment was carried to verifiy the approximations. This time, the relative error of $\hat{\alpha}$ is satisfying, and the estimation result is
			\begin{equation}
				\begin{aligned}
					\hat{\alpha} & = 0.100805\\
					\hat{\gamma} & = 0.198847
				\end{aligned}
			\end{equation}

			Further, we can apply the scheme described in part(e) without tuning  $E(n^2 x)$ finely, just take it as $E({n^*}^2 x^*)$, which also gives similiar result. But these trival stuff will not be discussed here as it has been clear enough for this problem...
	\end{enumerate}
